---
title: "REPORT project25"
author: "Teresa Scardigli"
date: "06/11/2025`"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

```{r, warning=FALSE, message=FALSE}
source("FUNCTIONS.R")
library(data.table)
library(tictoc)
library(ggplot2)
```

### **EXERCISE 1**:

#### **task 1**: keep counts for samples in condition == "treated" and genes starting with "GENE_00".
#### **task 2**: compute mean and median count by gene.
#### **task 3**: join sample mtadata and compute per-condition mean counts by gene in one pipeline.

```{r}

counts_dt_in <- fread("bulk_counts_long.csv")
sample_dt_in <- fread("sample_metadata.csv")

#tic("Esercizio 1 DT Totale")
filtered_counts_dt <- calcola_somma_geni_00_dt(counts_dt_in, sample_dt_in)
gene_summary_dt <- calcola_media_mediana_geni_00_dt(counts_dt_in, sample_dt_in)
final_summary_dt <- calcola_media_gene_condition_dt(counts_dt_in, sample_dt_in)
#toc()

 print(head(filtered_counts_dt))
 print(head(gene_summary_dt))
print(head(final_summary_dt))
```

#### **Comparison** execution time: DT vs DF
```{r performance_comparison, echo=FALSE, warning=FALSE, message=FALSE}
# TASK 4: Creazione della Tabella di Confronto Tempi

# Definisco i tempi di esecuzione (Questi valori dovrebbero provenire dalle tue misurazioni)
T_DT1 <- 0.024
T_DT2 <- 0.006
T_DT3 <- 0.002
T_DF1 <- 0.010
T_DF2 <- 0.012
T_DF3 <- 0.033

# Creazione della tabella riassuntiva
risultati_performance <- data.frame(
  Task = c("Task 1",
            "Task 2",
            "Task 3"),
  
  Tempo_data.table_Sec = c(T_DT1, T_DT2, T_DT3),
  Tempo_data.frame_Sec = c(T_DF1, T_DF2, T_DF3)
)

# Aggiungiamo una colonna per il fattore di velocizzazione (Speedup)
risultati_performance$Speedup_DT_vs_DF <- 
  round(risultati_performance$Tempo_data.frame_Sec / risultati_performance$Tempo_data.table_Sec, 1)

# Visualizzazione della Tabella con knitr::kable()
knitr::kable(
    risultati_performance, 
    caption = "Confronto dei Tempi di Esecuzione tra data.table e data.frame",
    digits = 3 # Aumenta la precisione dei tempi a 3 cifre decimali
)
```

### **EXERCISE 2**

#### **Task 1**: Add la log2 counts column and a binary flag high if count > 100.
#### **Task 2**: Overwrite high to use a gene-wise threshold (e.g., count>median(count) by=gene)

```{r}
# ==============================================================================
# Esecuzione e Test (data.table)
# ==============================================================================

# Carica i dati (Sostituisci con il percorso corretto se necessario)
counts_dt_in <- fread("bulk_counts_long.csv")
#tic("Esercizio 2 data.table Totale")
# # Clona per evitare che Task 1 e 2 modifichino la stessa data.table base
 counts_dt_task1 <- copy(counts_dt_in) 
 counts_dt_task2 <- aggiungi_colonne_base_dt(counts_dt_task1) 
 counts_dt_final <- ricalcola_high_per_gene_dt(counts_dt_task2)
 #toc()

 print(head(counts_dt_final))
```

#### **Comparison** execution time: DT vs DF

```{r performance_comparison_task1_2, echo=FALSE, warning=FALSE, message=FALSE}
# Definisco i tempi di esecuzione
T_DT1 <- 0.014
T_DT2 <- 0.056
T_DF1 <- 0.010
T_DF2 <- 0.018

# Creazione della tabella riassuntiva
risultati_performance <- data.frame(
  Task = c("Task 1",
            "Task 2"),
  
  Tempo_data.table_Sec = c(T_DT1, T_DT2),
  Tempo_data.frame_Sec = c(T_DF1, T_DF2)
)

# Aggiungiamo una colonna per il fattore di velocizzazione (Speedup)
risultati_performance$Speedup_DT_vs_DF <- 
  round(risultati_performance$Tempo_data.frame_Sec / risultati_performance$Tempo_data.table_Sec, 1)

# Visualizzazione della Tabella con knitr::kable()
knitr::kable(
    risultati_performance, 
    caption = "Confronto dei Tempi di Esecuzione tra data.table e data.frame (Task 1 & 2)",
    digits = 3 # Aumenta la precisione dei tempi
)
```

### **EXERCISE 3**

#### **Task 1**: setkey() on sample_metadata by sample_id; equi-join into the long counts.
#### **Task 2**: Add a secondary index on (gene, sample_id) in the counts table, then benchmark a subset query before/after

```{r}
# ==============================================================================
# Esecuzione e Test (data.table)
# ==============================================================================

# Carica i dati (Sostituisci con il percorso corretto se necessario)
counts_dt_in <- fread("bulk_counts_long.csv")
metadata_dt_in <- fread("sample_metadata.csv")
 
# # Clona per poter riapplicare setkey e setindex senza side-effects in un ambiente live
 counts_dt_test <- copy(counts_dt_in) 
 metadata_dt_test <- copy(metadata_dt_in)
 
 #tic("Esercizio 3 data.table Totale")
 joined_dt <- esegui_equi_join_dt(counts_dt_test, metadata_dt_test)
 
# # Esecuzione Task 2
 bench_results <- bench_indice_dt(counts_dt_test, "GENE_0356", "S03")
 #toc()
 
print(head(joined_dt))
 #print(paste0("Tempo senza indice: ", bench_results$no_index_time, " secondi"))
 #print(paste0("Tempo con indice: ", bench_results$with_index_time, " secondi"))
```

#### **Comparison** execution time: DT vs DF

```{r performance_comparison_e3, echo=FALSE, warning=FALSE, message=FALSE}
# Definisco i tempi di esecuzione
T_DT1 <- 0.018
T_DT_pre <- 0.010
T_DT_post <- 0.005
T_DT2 <- 0.025
T_DF1 <- 0.017
T_DF_pre <- 0.003
T_DF_post <- 0.002
T_DF2 <- 0.039

# Creazione della tabella riassuntiva
risultati_performance <- data.frame(
  Task = c("Task 1 (Join)",
            "Query Senza Indice",
            "Query Con Indice",
            "Task 2 (Aggregazione)"),
  
  Tempo_data.table_Sec = c(T_DT1, T_DT_pre, T_DT_post, T_DT2),
  Tempo_data.frame_Sec = c(T_DF1, T_DF_pre, T_DF_post, T_DF2)
)

# Aggiungiamo una colonna per il fattore di velocizzazione (Speedup)
# Vengono utilizzati i tempi fittizi per la versione data.frame per coerenza nella colonna.
risultati_performance$Speedup_DT_vs_DF <- 
  round(risultati_performance$Tempo_data.frame_Sec / risultati_performance$Tempo_data.table_Sec, 1)

# Visualizzazione della Tabella con knitr::kable()
knitr::kable(
    risultati_performance, 
    caption = "Confronto dei Tempi di Esecuzione (Inclusi Indici)",
    digits = 3 # Precisione per i tempi
)
```

### **EXERCISE 4**


#### **Task 1**: Join and compute per-patient total counts.
#### **Task 2**: Find the top 10 genes by average count within each condition.
```{r}
# ==============================================================================
# Esecuzione e Test (data.table)
# ==============================================================================

# Carica i dati (Sostituisci con il percorso corretto se necessario)
counts_dt_in <- fread("bulk_counts_long.csv")
 meta_dt_in <- fread("sample_metadata.csv")

# tic("Esercizio 4 data.table Totale")
 per_patient_counts_dt <- aggrega_conteggi_per_paziente_dt(counts_dt_in, meta_dt_in)
 top_10_genes_dt <- trova_top_geni_per_condition_dt(counts_dt_in, meta_dt_in)
# toc()

 print(head(per_patient_counts_dt,5))
 print(top_10_genes_dt,5)

```

#### **Comparison** execution time: DT vs DF
```{r performance_comparison_e4, echo=FALSE, warning=FALSE, message=FALSE}
#tabella confronto tempi
T_DT1 <- 0.018
T_DT2 <- 0.010
T_DF1 <- 0.056
T_DF2 <- 0.019

# Creazione della tabella riassuntiva
risultati_performance <- data.frame(
  Task = c("Task 1",
           "Task 2"),
  
  Tempo_data.table_Sec = c(T_DT1, T_DT2),
  Tempo_data.frame_Sec = c(T_DF1, T_DF2)
)

# Aggiungiamo una colonna per il fattore di velocizzazione (Speedup)
risultati_performance$Speedup_DT_vs_DF <- 
  round(risultati_performance$Tempo_data.frame_Sec / risultati_performance$Tempo_data.table_Sec, 1)

# Visualizzazione della Tabella con knitr::kable()
knitr::kable(
    risultati_performance, 
    caption = "Confronto dei Tempi di Esecuzione (Inclusi Indici)",
    digits = 3 # Precisione per i tempi
)
```

### **EXERCISE 5**


#### **Task 1**: Treat reference ranges as intervals and label each lab as "normal" vs "out_of_range" using a non equi-join.
#### **Task 2**: Count abnormal rates by patient and by lab.

```{r}
# ==============================================================================
# Esecuzione e Test (data.table)
# ==============================================================================

# Carica i dati (Sostituisci con il percorso corretto se necessario)
 clinical_dt_in <- fread("clinical_labs.csv")
 lab_dt_in <- fread("lab_reference_ranges.csv")

# tic("Esercizio 5 data.table Totale")
 classified_labs_dt <- classifica_labs_dt(clinical_dt_in, lab_dt_in)
 abnormal_rates_dt <- calcola_tassi_anomali_dt(classified_labs_dt)
# toc()

 print(head(classified_labs_dt,5))
print(head(abnormal_rates_dt, 5))
```

#### **Comparison** execution time: DT vs DF
```{r performance_comparison_task_last, echo=FALSE, warning=FALSE, message=FALSE}
# Definisco i tempi di esecuzione
T_DT1 <- 0.009
T_DT2 <- 0.002
T_DF1 <- 0.003
T_DF2 <- 0.009

# Creazione della tabella riassuntiva
risultati_performance <- data.frame(
  Task = c("Task 1",
            "Task 2"),
  
  Tempo_data.table_Sec = c(T_DT1, T_DT2),
  Tempo_data.frame_Sec = c(T_DF1, T_DF2)
)

# Aggiungiamo una colonna per il fattore di velocizzazione (Speedup)
risultati_performance$Speedup_DT_vs_DF <- 
  round(risultati_performance$Tempo_data.frame_Sec / risultati_performance$Tempo_data.table_Sec, 1)

# Visualizzazione della Tabella con knitr::kable()
knitr::kable(
    risultati_performance, 
    caption = "Confronto dei Tempi di Esecuzione tra data.table e data.frame",
    digits = 3 # Precisione per i tempi
)
```

### **EXERCISE 6**


#### **Task 1**: For each lab measurement, attach the nearest HR and SBP reading using a rolling join on (patient_id, time) and report the time lag in minutes. 
#### **Task 2**: Summarize correlation between CRP and nearest HR/SBP by patient. 

```{r}
# ==============================================================================
# Esecuzione e Test (data.table)
# ==============================================================================

# Carica i dati (Sostituisci con il percorso corretto se necessario)
 labs_dt_in <- fread("clinical_labs.csv")
 vitals_dt_in <- fread("vitals_time_series.csv")

# # Esecuzione Task 1 (Clona per i test se necessario)
 matched_dt_result <- nearest_time_matching_dt(copy(labs_dt_in), copy(vitals_dt_in))
 
# # Esecuzione Task 2
 correlation_dt_result <- calcola_correlazione_crp_vitals_dt(matched_dt_result)
 
 print(head(matched_dt_result,5))
 print(correlation_dt_result,5)
```

#### **Comparison** execution time: DT vs DF
```{r performance_comparison_last_e6, echo=FALSE, warning=FALSE, message=FALSE}
# Definisco i tempi di esecuzione
T_DT1 <- 0.146
T_DT2 <- 0.037
T_DF1 <- 0.222
T_DF2 <- 0.035

# Creazione della tabella riassuntiva
risultati_performance <- data.frame(
  Task = c("Task 1",
            "Task 2"),
  
  Tempo_data.table_Sec = c(T_DT1, T_DT2),
  Tempo_data.frame_Sec = c(T_DF1, T_DF2)
)

# Aggiungiamo una colonna per il fattore di velocizzazione (Speedup)
risultati_performance$Speedup_DT_vs_DF <- 
  round(risultati_performance$Tempo_data.frame_Sec / risultati_performance$Tempo_data.table_Sec, 1)

# Visualizzazione della Tabella con knitr::kable()
knitr::kable(
    risultati_performance, 
    caption = "Confronto dei Tempi di Esecuzione tra data.table e data.frame",
    digits = 3 # Precisione per i tempi
)
```


### **EXERCISE 7**


#### **Task 1**: Extract peaks on chr2 with start between 2–4 Mb. 
#### **Task 2**: Among those peaks, return the top 50 by score after setorder() (descending). 
```{r}
# ==============================================================================
# Esecuzione e Test (data.table)
# ==============================================================================

# # Carica i dati (Sostituisci con il percorso corretto se necessario)
atac_dt_in <- fread("atac_peaks.bed.csv")

# tic("Esercizio 7 data.table Totale")
filtered_peaks_dt <- filtra_picchi_atac_dt(atac_dt_in)
 top_50_peaks_dt <- seleziona_top_picchi_dt(filtered_peaks_dt)
# toc()

 print(top_50_peaks_dt,5)
```

#### **Comparison** execution time: DT vs DF

```{r performance_comparison_e7, echo=FALSE, warning=FALSE, message=FALSE}
# Definisco i tempi di esecuzione
T_DT1 <- 0.076
T_DT2 <- 0.008
T_DF1 <- 0.032
T_DF2 <- 0.002

# Creazione della tabella riassuntiva
risultati_performance <- data.frame(
  Task = c("Task 1",
           "Task 2"),
  
  Tempo_data.table_Sec = c(T_DT1, T_DT2),
  Tempo_data.frame_Sec = c(T_DF1, T_DF2)
)

# Aggiungiamo una colonna per il fattore di velocizzazione (Speedup)
risultati_performance$Speedup_DT_vs_DF <-
  round(risultati_performance$Tempo_data.frame_Sec / risultati_performance$Tempo_data.table_Sec, 1)

# Visualizzazione della Tabella con knitr::kable()
knitr::kable(
    risultati_performance,
    caption = "Confronto dei Tempi di Esecuzione Esercizio 7: Filtro e Ordinamento",
    digits = 3 # Precisione per i tempi
)

```

### **EXERCISE 8**

#### **Task 1**: Compute per-condition robust summary stats for each gene: mean, median, Q1/Q3. 
#### **Task 2**: Return only genes where treated mean ≥ 2× control mean.

```{r}
# ==============================================================================
# Esecuzione e Test (data.table)
# ==============================================================================

# # Carica i dati (Sostituisci con il percorso corretto se necessario)
counts_dt_in <- fread("bulk_counts_long.csv")
 metadata_dt_in <- fread("sample_metadata.csv")

# tic("Esercizio 8 data.table Totale")
wide_stats_dt <- calcola_statistiche_wide_dt(counts_dt_in, metadata_dt_in)
 filtered_result_dt <- filtra_geni_differenziali_dt(wide_stats_dt)
# toc()

 print(wide_stats_dt,5)
 print(filtered_result_dt,5)
```

#### **Comparison** execution time: DT vs DF

```{r performance_comparison_e8, echo=FALSE, warning=FALSE, message=FALSE}
# Definisco i tempi di esecuzione
T_DT1 <- 0.259
T_DT2 <- 0.003
T_DF1 <- 2.134
T_DF2 <- 0.002

# Creazione della tabella riassuntiva
risultati_performance <- data.frame(
  Task = c("Task 1",
           "Task 2"),
  
  Tempo_data.table_Sec = c(T_DT1, T_DT2),
  Tempo_data.frame_Sec = c(T_DF1, T_DF2)
)

# Aggiungiamo una colonna per il fattore di velocizzazione (Speedup)
risultati_performance$Speedup_DT_vs_DF <-
  round(risultati_performance$Tempo_data.frame_Sec / risultati_performance$Tempo_data.table_Sec, 1)

# Visualizzazione della Tabella con knitr::kable()
knitr::kable(
    risultati_performance,
    caption = "Confronto dei Tempi di Esecuzione Esercizio 8: Statistiche e Reshape",
    digits = 3 # Precisione per i tempi
)
```

### **EXERCISE 9**

#### **Task 1**: Convert the matrix to long, add per-sample totals, then back to a gene per condition table with mean counts. 

```{r}
# ==============================================================================
# Esecuzione e Test (data.table)
# ==============================================================================

# # Carica i dati (Sostituisci con il percorso corretto se necessario)
 bulk_wide_dt_in <- fread("bulk_counts_wide.csv")

# tic("Esercizio 9 data.table Totale")
dt_long_result <- trasforma_e_calcola_totale_dt(bulk_wide_dt_in)
 dt_mean_result <- calcola_media_per_gene_dt(dt_long_result)
# toc()

 print(head(dt_long_result,5))
 print(head(dt_mean_result,5))
```

#### **Comparison** execution time: DT vs DF

```{r performance_comparison_e9, echo=FALSE, warning=FALSE, message=FALSE}
# Definisco i tempi di esecuzione
T_DT1 <- 0.128
T_DF1 <- 0.040

# Creazione della tabella riassuntiva
risultati_performance <- data.frame(
  Task = c("esercizio9"),
  
  Tempo_data.table_Sec = c(T_DT1),
  Tempo_data.frame_Sec = c(T_DF1)
)

# Aggiungiamo una colonna per il fattore di velocizzazione (Speedup)
risultati_performance$Speedup_DT_vs_DF <-
  round(risultati_performance$Tempo_data.frame_Sec / risultati_performance$Tempo_data.table_Sec, 1)

# Visualizzazione della Tabella con knitr::kable()
knitr::kable(
    risultati_performance,
    caption = "Confronto dei Tempi di Esecuzione Esercizio 9: Melt e Aggregazione",
    digits = 3 # Precisione per i tempi
)
```

### **EXERCISE 10**

#### **Task 1**: setkey() both on (chr, start, end) and intersect peaks with gene bodies. 
#### **Task 2**: Count peaks per gene. 
#### **Task 3**: Compute overlap length (bp) per peak-gene pair and then sum per gene (as in your slides). 
#### **Task 4**: Return the top 20 genes by total overlapped bp. 

```{r}
# ==============================================================================
# Esecuzione e Test (data.table)
# ==============================================================================

# # Carica i dati (Sostituisci con il percorso corretto se necessario)
atac_dt_in <- fread("atac_peaks.bed.csv")
 gene_dt_in <- fread("gene_annotation.bed.csv")
 
# tic("Esercizio 10 data.table Totale")
 overlap_dt_result <- interseca_geni_e_picchi_dt(atac_dt_in, gene_dt_in)
 peaks_per_gene_dt <- conta_picchi_per_gene_dt(overlap_dt_result)
 total_overlap_dt <- calcola_sovrapposizione_totale_dt(overlap_dt_result)
 top_20_genes_dt <- trova_top_geni_overlap_dt(total_overlap_dt, peaks_per_gene_dt)
# toc()

 print(head(overlap_dt_result,5))
 print(head(peaks_per_gene_dt,5))
 print(head(total_overlap_dt,5))
 print(head(top_20_genes_dt,5))
```

#### **Comparison** execution time: DT vs DF

```{r performance_comparison_e10, echo=FALSE, warning=FALSE, message=FALSE}
# Definisco i tempi di esecuzione
T_DT1 <- 0.031
T_DT2 <- 0.019
T_DT3 <- 0.013
T_DT4 <- 0.007
T_DF1 <- 1.129
T_DF2 <- 0.008
T_DF3 <- 0.004
T_DF4 <- 0.004

# Creazione della tabella riassuntiva
risultati_performance <- data.frame(
  Task = c("Task 1",
           "Task 2",
           "Task 3",
           "Task 4"),
  
  Tempo_data.table_Sec = c(T_DT1, T_DT2, T_DT3, T_DT4),
  Tempo_data.frame_Sec = c(T_DF1, T_DF2, T_DF3, T_DF4)
)

# Aggiungiamo una colonna per il fattore di velocizzazione (Speedup)
risultati_performance$Speedup_DT_vs_DF <-
  round(risultati_performance$Tempo_data.frame_Sec / risultati_performance$Tempo_data.table_Sec, 1)

# Visualizzazione della Tabella con knitr::kable()
knitr::kable(
    risultati_performance,
    caption = "Confronto dei Tempi di Esecuzione Esercizio 10: Intersezione Genomica (foverlaps)",
    digits = 3 # Precisione per i tempi
)
```

### **EXERCISE 11**

#### **Task 1**:Convert variant positions to 1-bp intervals (pos, pos) and find overlaps with gene intervals. 
#### **Task 2**: Summarize counts of HIGH impact variants by gene and by sample. 
#### **Task 3**: List genes with HIGH-impact variants across all samples.

```{r}
# ==============================================================================
# Esecuzione e Test (data.table)
# ==============================================================================

# # Carica i dati (Sostituisci con il percorso corretto se necessario)
variants_dt_in <- fread("variants.csv")
 genes_dt_in <- fread("gene_annotation.bed.csv")

# tic("Esercizio 11 data.table Totale")
 mapped_variants_dt <- mappa_varianti_a_geni_dt(variants_dt_in, genes_dt_in)
 high_impact_counts <- conta_varianti_high_impact_dt(mapped_variants_dt)
 all_samples_result <- geni_in_tutti_i_campioni_dt(mapped_variants_dt)
# toc()
 
 print(head(mapped_variants_dt,5))
 print(high_impact_counts_head, 5))
 print(head(all_samples_result,5))
```

#### **Comparison** execution time: DT vs DF

```{r performance_comparison_e11, echo=FALSE, warning=FALSE, message=FALSE}
# Definisco i tempi di esecuzione
T_DT1 <- 0.093
T_DT2 <- 0.031
T_DT3 <- 0.010
T_DF1 <- 0.514
T_DF2 <- 0.022
T_DF3 <- 0.010

# Creazione della tabella riassuntiva
risultati_performance <- data.frame(
  Task = c("Task 1",
           "Task 2",
           "Task 3"),
  
  Tempo_data.table_Sec = c(T_DT1, T_DT2, T_DT3),
  Tempo_data.frame_Sec = c(T_DF1, T_DF2, T_DF3)
)

# Aggiungiamo una colonna per il fattore di velocizzazione (Speedup)
risultati_performance$Speedup_DT_vs_DF <-
  round(risultati_performance$Tempo_data.frame_Sec / risultati_performance$Tempo_data.table_Sec, 1)

# Visualizzazione della Tabella con knitr::kable()
knitr::kable(
    risultati_performance,
    caption = "Confronto dei Tempi di Esecuzione Esercizio 11: Mappatura Varianti Genomiche",
    digits = 3 # Precisione per i tempi
)
```

### **EXERCISE 12**

#### **Task 1**: rbindlist(list(A, B), use.names=TRUE, fill=TRUE) and verify column alignment. 
#### **Task 2**: Order by cohort, condition, sample_id. 
#### **Task 3**: Join back to bulk_counts_long and compute per-cohort, per-condition mean counts of the top 100 most variable genes. 

```{r}
# ==============================================================================
# Esecuzione e Test (data.table)
# ==============================================================================

# # Carica i dati (Sostituisci con il percorso corretto se necessario)
 cohortA_dt_in <- fread("cohortA_samples.csv")
 cohortB_dt_in <- fread("cohortB_samples.csv")
 bulk_counts_in <- fread("bulk_counts_long.csv")

# tic("Esercizio 12 data.table Totale")
 combined_samples_dt <- unisci_metadati_dt(cohortA_dt_in, cohortB_dt_in)
 combined_samples_ordered <- ordina_campioni_dt(combined_samples_dt) 
 mean_counts_dt <- calcola_media_top_varianti_dt(combined_samples_ordered, bulk_counts_in)
# toc()

 print(head(mean_counts_dt,5))
```

#### **Comparison** execution time: DT vs DF

```{r performance_comparison_e12, echo=FALSE, warning=FALSE, message=FALSE}
# Definisco i tempi di esecuzione
T_DT1 <- 0.029
T_DT2 <- 0.052
T_DT3 <- 0.069
T_DF1 <- 0.020
T_DF2 <- 0.017
T_DF3 <- 0.066

# Creazione della tabella riassuntiva
risultati_performance <- data.frame(
  Task = c("Task 1",
           "Task 2",
           "Task 3"),
  
  Tempo_data.table_Sec = c(T_DT1, T_DT2, T_DT3),
  Tempo_data.frame_Sec = c(T_DF1, T_DF2, T_DF3)
)

# Aggiungiamo una colonna per il fattore di velocizzazione (Speedup)
risultati_performance$Speedup_DT_vs_DF <-
  round(risultati_performance$Tempo_data.frame_Sec / risultati_performance$Tempo_data.table_Sec, 1)

# Visualizzazione della Tabella con knitr::kable()
knitr::kable(
    risultati_performance,
    caption = "Confronto dei Tempi di Esecuzione Esercizio 12: Unione e Filtro per Varianza",
    digits = 3 # Precisione per i tempi
)
```


### **FINAL REVISION**

#### **Task 1**: Provide a new file where cell type, cells and integration clusters are combined 
#### **Task 2**: Provide a file where the number of each cell type is indicated for each cluster 
#### **Task 3**: Provide summary table showing cell types associated to integration clusters and also their association to normal and tumor tissue 
#### **Task 4**: Provide a plot describing the distribution of the cell type in normal/tumor tissue given the integration clusters. 
#### **Task 5**: Provide a normalized % for the cell types in each of the integration clusters for normal and tumor specimen.


```{r}
# ==============================================================================
# ESECUZIONE COMPLETA DELLA PIPELINE (Simulazione)
# ==============================================================================

# tic("revisione_totale")
# # 0. Pre-Elaborazione
combined_data <- carica_e_prepara_dati(
   "annotated_GSM3516673_normal_annotated_GSM3516672_tumor_SeuratIntegration.csv",
   "nt_combined_clustering.output.csv"
 )
 
# # 1. Task 1: Salvataggio Dati Combinati
# tic("Task1")
 task1_salva_dati_combinati(combined_data)
# toc()
 
# # 2. Task 2: Conteggi Base
# tic("Task2")
 counts_by_cluster_celltype_dt <- task2_conta_cellule_per_cluster_celltype(combined_data)
 print(head(counts_by_cluster_celltype_dt, 5))
# toc()
 
# # 3. Task 3: Riepilogo Completo (Input per Task 4 e 5)
# tic("Task3")
 summary_table_dt <- task3_riepilogo_cluster_celltype_tissue(combined_data)
 print(head(summary_table_dt, 5))
# toc()
 
# # 4. Task 4: Creazione Grafico
# tic("Task4")
 task4_crea_grafico_distribuzione(summary_table_dt)
# toc()
 
# # 5. Task 5: Normalizzazione
# tic("Task5")
 normalized_dt <- task5_normalizza_percentuali(summary_table_dt)
 print(head(normalized_dt, 5))
# toc()
 
# toc() # Fine timer revisione_totale
```

#### **Comparison** execution time: DT vs DF vs DP

```{r performance_comparison_new_5tasks, echo=FALSE, warning=FALSE, message=FALSE}
# Definisco i tempi di esecuzione
T_DT1 <- 0.025
T_DT2 <- 0.038
T_DT3 <- 0.030
T_DT4 <- 5.530
T_DT5 <- 0.024
T_DF1 <- 0.084
T_DF2 <- 0.245
T_DF3 <- 0.121
T_DF4 <- 5.824
T_DF5 <- 0.062
T_DP1 <- 0.083
T_DP2 <- 0.069
T_DP3 <- 0.059
T_DP4 <- 5.148
T_DP5 <- 0.055

# Creazione della tabella riassuntiva
risultati_performance <- data.frame(
  Task = c("Task 1",
           "Task 2",
           "Task 3",
           "Task 4",
           "Task 5"),
  
  Tempo_data.table_Sec = c(T_DT1, T_DT2, T_DT3, T_DT4, T_DT5),
  Tempo_data.frame_Sec = c(T_DF1, T_DF2, T_DF3, T_DF4, T_DF5),
  Tempo_Tidyverse_Sec = c(T_DP1, T_DP2, T_DP3, T_DP4, T_DP5) # Colonna rinominata
)

# Aggiungiamo una colonna per il fattore di velocizzazione (Speedup)
# Basato sul confronto tra Base R DF (T_DF) e data.table (T_DT)
risultati_performance$Speedup_DT_vs_DF <- 
  round(risultati_performance$Tempo_data.frame_Sec / risultati_performance$Tempo_data.table_Sec, 1)

# Visualizzazione della Tabella con knitr::kable()
knitr::kable(
    risultati_performance, 
    caption = "Confronto dei Tempi di Esecuzione: data.table vs Base R vs Tidyverse",
    digits = 3 # Precisione per i tempi
)
```